Realizability-Preserving DG-IMEX Method for the Two-Moment Model of Fermion Transport \tnoteref{support}

### 2018-09-01 Documents/M1/realizableFermionicM1/sections/imex.tex
Edits througout after reading the entire paper.

Eq.~\eqref{eq:ode} is to be integrated forward in time with an ODE integrator, and we begin with the following definition of a constraint-preserving time integration scheme, as applied to integrate a physical model forward in time.  
\begin{define}
  Given sufficient conditions, a constraint-preserving time integration scheme preserves the constraints of a model if the set of admissible states satisfying the constraints form a convex set.  
  \label{def:constraintPreserving}
\end{define}
As an example, given a spatial discretization of a model that is constraint-preserving with the forward Euler method for a sufficiently small time step, high-order strong stability-preserving Runge-Kutta (SSP-RK) methods, which can be expressed as convex combinations of forward Euler steps, form a class of constraint-preserving schemes.

@@

Unfortunately, the ``diffusion limit requirement" in Eq.~\eqref{eq:diffusionCondition}, together with the order conditions given by Eqs.~\eqref{orderConditions1} and \eqref{orderConditions2} (with $\alpha=0$), and the positivity conditions on $c_{ij}$ and $\tilde{c}_{ij}$, result in too many constraints to obtain a constraint-preserving IMEX scheme.
\footnote{It was shown in \cite{hu_etal_2018} --- without the diffusion limit requirement --- that the minimum number of stages for constraint-preserving IMEX schemes of type~ARS is four.)  
We are also concerned about increasing the number of stages, and thereby the number of implicit solves, since the implicit solve will dominate the computational cost of the IMEX scheme with more realistic collision operators (e.g., inelastic scattering).}  
To reduce the number of constraints, and accommodate accuracy in the diffusion limit, we relax the requirement of overall second-order accuracy of the IMEX scheme.  
Instead, we only require the scheme to be second-order accurate in the streaming limit ($\vect{\cQ}=0$).  
This gives the order conditions
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}=1
  \quad\text{and}\quad
  \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}=\f{1}{2},
  \label{eq:orderConditionsEx}
\end{equation}
where the first condition (consistency condition) is required for first-order accuracy.  
We then seek to design IMEX schemes of Type~ARS conforming to the following working definition
\begin{define}
  Let PD-IMEX be a consistent, constraint-preserving, diffusion accurate, and low-stage GSA IMEX scheme satisfying the following properties
  \begin{enumerate}
    \item Consistency of the implicit coefficients
    \begin{equation}
      \sum_{i=1}^{s}w_{i}=1.
      \label{eq:implicitConsistency}
    \end{equation}
    \item Second-order in the streaming limit; i.e., satisfies Eq.~\eqref{eq:orderConditionsEx}.
    \item Constraint-preserving; i.e. satisfies Eq.~\eqref{eq:positivityConditionsTypeARS}, with $\sum_{j=0}^{i-1}c_{ij}=1$, for $i=1,\ldots,s$, and $c_{\mbox{\tiny Sch}}>0$.
    \item Well-behaved in the diffusion limit; i.e., satisfies Eq.~\eqref{eq:diffusionCondition}.
    \item Less than four stages ($s\le3$).
    \item Globally stiffly accurate (GSA) $a_{si}=w_{i}$ and $\tilde{a}_{si}=\tilde{w}_{i},\quad i=1,\ldots,s$.
  \end{enumerate}
  \label{def:PD-IMEX}
\end{define}
Fortunately, IMEX schemes of type~ARS satisfying these properties are easy to find, and we provide an example with $s=3$ (two implicit solves) in \ref{app:butcherTables} (scheme PD-ARS; see \ref{app:PD-ARS} for details).  
In the streaming limit, this scheme is identical to the optimal second-order accurate SSP-RK method \cite{gottlieb_etal_2001}.  
It is also very similar to the scheme given in \cite{mcclarren_etal_2008} (see scheme PC2 in \ref{app:butcherTables}), which is also a GSA IMEX scheme of type ARS with $s=3$.  
Scheme PC2 is second-order in the streaming limit, has been demonstrated to work well in the diffusion limit \cite{mcclarren_etal_2008,radice_etal_2013}, and satisfies the positivity conditions.  
However, $c_{\mbox{\tiny Sch}}=0$ (our primary motivation for finding an alternative).  
In Section~\ref{sec:numerical}, we show numerically that the accuracy of scheme PD-ARS is comparable to the accuracy of scheme PC2.

@@

Here we analyze the absolute stability of the proposed IMEX schemes, PA2+ and PD-ARS (given in \ref{app:butcherTables}), following \cite{hu_etal_2018}.  
As is commonly done, we do this in the context of the linear scalar equation
\begin{equation}
  \dot{u}=\lambda_{1}\,u+\lambda_{2}\,u,
  \label{eq:scalarODE}
\end{equation}
where $\lambda_{1}\in\mathbb{C}$ and real $\lambda_{2}\le0$.  
On the right-hand side of Eq.~\eqref{eq:scalarODE}, the first (oscillatory) term is treated explicitly, while the second (damping) term is treated implicitly.  
The IMEX schemes can then be written as $u^{n+1} =P(z_{1},z_{2})\,u^{n}$, where $P(z_{1},z_{2})$ is the amplification factor of the scheme, $z_{1}=\dt\,\lambda_{1} = x +iy$, and $z_{2} = \dt\,\lambda_{2} < 0$.  
Stability of the IMEX scheme requires $|P(z_1,z_2)|\leq 1$.  
The stability regions of PA2+ and PD-ARS are plotted in Figure~\ref{fig:AbsoluteStability}.  
As can be seen from the figures, the absolute stability region of both schemes increases with increasing $|z_{2}|$ (increased damping).  
For a given $|z_{2}|>0$, the stability region of PD-ARS is larger than that of PA2+.  
In the linear case, a time step that satisfies the absolute stability for the explicit part of the IMEX scheme ($|z_2| = 0$) fulfills the stability requirement for the IMEX scheme as whole ($|z_2| \geq 0$).  
This holds for all PD-ARS schemes with $\epsilon\in [0,0.5)$.
\begin{figure}[h]
  \centering
  \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{figures/AbsoluteStabilityPARSD}
    \includegraphics[width=0.5\textwidth]{figures/AbsoluteStabilityPA2+}
  \end{tabular}
   \caption{Boundary of the absolute stability region for different values of $z_2$ for PD-ARS with $\epsilon = 0.1$ (left panel) and PA2+ (right panel).}
  \label{fig:AbsoluteStability}
\end{figure}

>>>>

Eq.~\eqref{eq:ode} is to be integrated forward in time with an ODE integrator, and we begin with the following definition of a constraint-preserving time integration scheme, as applied to integrate a physical model forward in time.  
\begin{define}
  Given sufficient conditions, a constraint-preserving time integration scheme preserves the constraints of a model if the set of admissible states satisfying the constraints form a convex set.  
  \label{def:constraintPreserving}
\end{define}
As an example, given a spatial discretization of a model that is constraint-preserving with the forward Euler method for a sufficiently small time step, high-order explicit strong stability-preserving Runge-Kutta (SSP-RK) methods, which can be expressed as convex combinations of forward Euler steps, also preserve the constraints (possibly with a smaller time step), and form a class of constraint-preserving schemes.

Since the realizable set $\cR$ is convex, constraint-preserving schemes can be used to design realizability-preserving schemes for the two-moment model.

@@

Unfortunately, the ``diffusion limit requirement" in Eq.~\eqref{eq:diffusionCondition}, together with the order conditions given by Eqs.~\eqref{orderConditions1} and \eqref{orderConditions2} (with $\alpha=0$), and the positivity conditions on $c_{ij}$ and $\tilde{c}_{ij}$, result in too many constraints to obtain a second-order accurate constraint-preserving IMEX scheme.
\footnote{It was shown in \cite{hu_etal_2018} --- without the diffusion limit requirement --- that the minimum number of stages for constraint-preserving IMEX schemes of type~ARS is four.)  
We are also concerned about increasing the number of stages, and thereby the number of implicit solves, since the implicit solve will dominate the computational cost of the IMEX scheme with more realistic collision operators (e.g., inelastic scattering).}  
To reduce the number of constraints, and accommodate accuracy in the diffusion limit, we relax the requirement of overall second-order accuracy of the IMEX scheme.  
Instead, we only require the scheme to be second-order accurate in the streaming limit ($\vect{\cQ}=0$).  
This gives the order conditions
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}=1
  \quad\text{and}\quad
  \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}=\f{1}{2},
  \label{eq:orderConditionsEx}
\end{equation}
where the first condition (consistency condition) is required for first-order accuracy.  
We then seek to design IMEX schemes of Type~ARS conforming to the following working definition
\begin{define}
  Let PD-IMEX be a consistent, constraint-preserving, diffusion accurate, and low-stage GSA IMEX scheme satisfying the following properties
  \begin{enumerate}
    \item Consistency of the implicit coefficients
    \begin{equation}
      \sum_{i=1}^{s}w_{i}=1.
      \label{eq:implicitConsistency}
    \end{equation}
    \item Second-order in the streaming limit; i.e., satisfies Eq.~\eqref{eq:orderConditionsEx}.
    \item Constraint-preserving; i.e. satisfies Eq.~\eqref{eq:positivityConditionsTypeARS}, with $\sum_{j=0}^{i-1}c_{ij}=1$, for $i=1,\ldots,s$, and $c_{\mbox{\tiny Sch}}>0$.
    \item Well-behaved in the diffusion limit; i.e., satisfies Eq.~\eqref{eq:diffusionCondition}.
    \item Less than four stages ($s\le3$).
    \item Globally stiffly accurate (GSA) $a_{si}=w_{i}$ and $\tilde{a}_{si}=\tilde{w}_{i},\quad i=1,\ldots,s$.
  \end{enumerate}
  \label{def:PD-IMEX}
\end{define}
Fortunately, IMEX schemes of type~ARS satisfying these properties are easy to find, and we provide an example with $s=3$ (two implicit solves) in \ref{app:butcherTables} (scheme PD-ARS; see \ref{app:PD-ARS} for further details).  
In the streaming limit, this scheme is identical to the optimal second-order accurate SSP-RK method \cite{gottlieb_etal_2001}.  
It is also very similar to the scheme given in \cite{mcclarren_etal_2008} (see scheme PC2 in \ref{app:butcherTables}), which is also a GSA IMEX scheme of type ARS with $s=3$.  
Scheme PC2 is second-order in the streaming limit, has been demonstrated to work well in the diffusion limit \cite{mcclarren_etal_2008,radice_etal_2013}, and satisfies the positivity conditions in Eq.~\eqref{eq:positivityConditionsTypeARS}.  
However, $c_{\mbox{\tiny Sch}}=0$ (our primary motivation for finding an alternative).  
In Section~\ref{sec:numerical}, we show numerically that the accuracy of scheme PD-ARS is comparable to the accuracy of scheme PC2.

@@

Here we analyze the absolute stability of the proposed IMEX schemes, PA2+ and PD-ARS (given in \ref{app:butcherTables}), following \cite{hu_etal_2018}.  
As is commonly done, we do this in the context of the linear scalar equation
\begin{equation}
  \dot{u}=\lambda_{1}\,u+\lambda_{2}\,u,
  \label{eq:scalarODE}
\end{equation}
where $\lambda_{1}\in\mathbb{C}$ and real $\lambda_{2}\le0$.  
On the right-hand side of Eq.~\eqref{eq:scalarODE}, the first (oscillatory) term is treated explicitly, while the second (damping) term is treated implicitly.  
The IMEX schemes can then be written as $u^{n+1} =P(z_{1},z_{2})\,u^{n}$, where $P(z_{1},z_{2})$ is the amplification factor of the scheme, $z_{1}=\dt\,\lambda_{1} = x +iy$, and $z_{2} = \dt\,\lambda_{2} \le 0$.  
Stability of the IMEX scheme requires $|P(z_1,z_2)|\leq 1$.  
The stability regions of PA2+ and PD-ARS are plotted in Figure~\ref{fig:AbsoluteStability}.  
As can be seen from the figures, the absolute stability region of both schemes increases with increasing $|z_{2}|$ (increased damping).  
For a given $|z_{2}|\ge0$, the stability region of PD-ARS is larger than that of PA2+.  
In the linear case, a time step that satisfies the absolute stability for the explicit part of the IMEX scheme ($|z_2| = 0$) fulfills the stability requirement for the IMEX scheme as whole ($|z_2| \geq 0$).  
This holds for all PD-ARS schemes with $\epsilon\in [0,0.5)$.
\begin{figure}[h]
  \centering
  \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{figures/AbsoluteStabilityPARSD}
    \includegraphics[width=0.5\textwidth]{figures/AbsoluteStabilityPA2+}
  \end{tabular}
   \caption{Boundary of the absolute stability region for different values of $z_2$ for PD-ARS with $\epsilon = 0.1$ (left panel) and PA2+ (right panel).}
  \label{fig:AbsoluteStability}
\end{figure}

### 2018-09-01 Documents/M1/realizableFermionicM1/sections/intro.tex
Edits througout after reading the entire paper.

In many applications, the particle mean free path is comparable to or exceeds other characteristic length scales in the system under consideration (i.e., the Knudsen number exceeds unity), and non-equilibrium effects may become important.  
In these situations, a kinetic description may be required, where particle distribution function $f(\vect{p},\vect{x},t)$ is solved for.  
The distribution function, a phase space density depending on momentum $\vect{p}$ and position $\vect{x}$, gives at time $t$ the number of particles in the phase space volume element $d\vect{p}\,d\vect{x}$ (i.e., $dN=f\,d\vect{p}\,d\vect{x}$).  
The evolution of the distribution function is governed by the Boltzmann equation, which states a balance between phase space advection and collisions (see, e.g., \cite{braginskii_1965,chapmanCowling_1970,lifshitzPitaevskii_1981}).

@@

In this paper we consider a two-moment model for particles governed by Fermi-Dirac statistics.  
It is well known from the two-moment model for particles governed by Maxwell-Boltzmann statistics (``classical'' particles with $f\ge0$), that the particle density is nonnegative and the magnitude of the flux vector is bounded by the particle density.  
(There are further constraints on the components of the Eddington tensor \cite{levermore_1984}.)  
Furthermore, the set of realizable moments generated by the particle density and flux vector constitutes a convex cone \cite{olbrant_etal_2012}.  
In the fermionic case, there is also an upper bound on the distribution function because Pauli's exclusion principle prevents particles from occupying the same microscopic state (e.g., $f\le1$).  
The fermionic two-moment model has recently been studied theoretically in the context of maximum entropy closures \cite{lareckiBanach_2011,banachLarecki_2013,banachLarecki_2017b} and Kershaw-type closures \cite{banachLarecki_2017a}.  
Because of the upper bound on the distribution function, the algebraic constraints on realizable moments differ from the classical case with no upper bound, and can lead to significantly different dynamics when the occupancy is high (i.e., when $f$ is close to its upper bound).  
In the fermionic case, the set of realizable moments generated by the particle density and flux vector is also convex, and resembles a ``pinched'' cone.  
It coincides with the realizable set for the classical case when the occupancy is low, but is much more restricted for high occupancy.

The two-moment model is discretized in space using high-order Discontinuous Galerkin (DG) methods (e.g., \cite{cockburnShu_2001,hesthavenWarburton_2008}).  
DG methods combine elements from both spectral and finite volume methods, and are an attractive option for solving hyperbolic partial differential equations (PDEs).  
They achieve high-order accuracy on a compact stencil; i.e., data is only communicated with nearest neighbors, regardless of the formal order of accuracy, which can lead to a high computation to communication ratio, and favorable parallel scalability on heterogeneous architectures has been demonstrated \cite{klockner_etal_2009}.  
Furthermore, they can easily be applied to problems involving curvilinear coordinates (e.g., beneficial in numerical relativity \cite{teukolsky_2016}).  
Importantly, DG methods, exhibit favorable properties when collisions with a background are included as they recover the correct asymptotic behavior in the diffusion limit, characterized by frequent collisions (e.g., \cite{larsenMorel_1989,adams_2001,guermondKanschat_2010}).  
The DG method was introduced in the 1970s by Reed \& Hill \cite{reedHill_1973} to solve the neutron transport equation, and has undergone remarkable developments since then (see, e.g., \cite{shu_2016}, and references therein).

We are concerned with the development and application of DG methods for the fermionic two-moment model that can preserve the aforementioned algebraic constraints and ensure realizable moments, provided the initial condition is realizable.  
Our approach is based on the constraint-preserving (CP) framework introduced in \cite{zhangShu_2010a}, and later extended to the Euler equations of gas dynamics in \cite{zhangShu_2010b}.  
(See e.g., \cite{xing_etal_2010,zhangShu_2011,olbrant_etal_2012,cheng_etal_2013,zhang_etal_2013,endeve_etal_2015,wuTang_2015}, for extensions and applications to other systems.)  
The main ingredients include (1) a realizability-preserving update for the cell averaged moments based on Euler time stepping, which evaluates the polynomial representation of the DG method in a finite number of quadrature points in the local elements and results in a Courant-Friedrichs-Lewy (CFL) condition on the time step; (2) a limiter to modify the polynomial representation to ensure that the algebraic constraints are satisfied point-wise without changing the cell average of the moments; and (3) a time stepping method that can be expressed as a convex combination of Euler steps, and therefore preserves the algebraic constraints as the Euler methods.

@@

\clearpage

>>>>

In many applications, the particle mean free path is comparable to or exceeds other characteristic length scales in the system under consideration (i.e., the Knudsen number exceeds unity), and non-equilibrium effects may become important.  
In these situations, a kinetic description may be required, where particle distribution function $f(\vect{p},\vect{x},t)$ is solved for.  
The distribution function, a phase space density depending on momentum $\vect{p}\in\bbR^{3}$ and position $\vect{x}\in\bbR^{3}$, gives at time $t\in\bbR^{+}$ the number of particles in the phase space volume element $d\vect{p}\,d\vect{x}$ (i.e., $dN=f\,d\vect{p}\,d\vect{x}$).  
The evolution of the distribution function is governed by the Boltzmann equation, which states a balance between phase space advection and collisions (see, e.g., \cite{braginskii_1965,chapmanCowling_1970,lifshitzPitaevskii_1981}).

@@

Here we consider a two-moment model for particles governed by Fermi-Dirac statistics.  
It is well known from the two-moment model for particles governed by Maxwell-Boltzmann statistics (``classical'' particles with $f\ge0$), that the particle density is nonnegative and the magnitude of the flux vector is bounded by the particle density.  
(There are further constraints on the components of the Eddington tensor \cite{levermore_1984}.)  
Furthermore, the set of realizable moments generated by the particle density and flux vector constitutes a convex cone \cite{olbrant_etal_2012}.  
In the fermionic case, there is also an upper bound on the distribution function (e.g., $f\le1$) because Pauli's exclusion principle prevents particles from occupying the same microscopic state.  
The fermionic two-moment model has recently been studied theoretically in the context of maximum entropy closures \cite{lareckiBanach_2011,banachLarecki_2013,banachLarecki_2017b} and Kershaw-type closures \cite{banachLarecki_2017a}.  
Because of the upper bound on the distribution function, the algebraic constraints on realizable moments differ from the classical case with no upper bound, and can lead to significantly different dynamics when the occupancy is high (i.e., when $f$ is close to its upper bound).  
In the fermionic case, the set of realizable moments generated by the particle density and flux vector is also convex, and resembles a ``pinched'' cone.  
It coincides with the realizable set for the classical case when the occupancy is low, but is much more restricted for high occupancy.

In this paper, the two-moment model is discretized in space using high-order Discontinuous Galerkin (DG) methods (e.g., \cite{cockburnShu_2001,hesthavenWarburton_2008}).  
DG methods combine elements from both spectral and finite volume methods, and are an attractive option for solving hyperbolic partial differential equations (PDEs).  
They achieve high-order accuracy on a compact stencil; i.e., data is only communicated with nearest neighbors, regardless of the formal order of accuracy, which can lead to a high computation to communication ratio, and favorable parallel scalability on heterogeneous architectures has been demonstrated \cite{klockner_etal_2009}.  
Furthermore, they can easily be applied to problems involving curvilinear coordinates (e.g., beneficial in numerical relativity \cite{teukolsky_2016}).  
Importantly, DG methods, exhibit favorable properties when collisions with a background are included as they recover the correct asymptotic behavior in the diffusion limit, characterized by frequent collisions (e.g., \cite{larsenMorel_1989,adams_2001,guermondKanschat_2010}).  
The DG method was introduced in the 1970s by Reed \& Hill \cite{reedHill_1973} to solve the neutron transport equation, and has undergone remarkable developments since then (see, e.g., \cite{shu_2016}, and references therein).

We are concerned with the development and application of DG methods for the fermionic two-moment model that can preserve the aforementioned algebraic constraints and ensure realizable moments, provided the initial condition is realizable.  
Our approach is based on the constraint-preserving (CP) framework introduced in \cite{zhangShu_2010a}, and later extended to the Euler equations of gas dynamics in \cite{zhangShu_2010b}.  
(See e.g., \cite{xing_etal_2010,zhangShu_2011,olbrant_etal_2012,cheng_etal_2013,zhang_etal_2013,endeve_etal_2015,wuTang_2015}, for extensions and applications to other systems.)  
The main ingredients include (1) a realizability-preserving update for the cell averaged moments based on Euler time stepping, which evaluates the polynomial representation of the DG method in a finite number of quadrature points in the local elements and results in a Courant-Friedrichs-Lewy (CFL) condition on the time step; (2) a limiter to modify the polynomial representation to ensure that the algebraic constraints are satisfied point-wise without changing the cell average of the moments; and (3) a time stepping method that can be expressed as a convex combination of Euler steps, and therefore preserves the algebraic constraints as the Euler methods (possibly with a modified CFL condition).

@@

### 2018-09-01 Documents/M1/realizableFermionicM1/sections/limiter.tex
Completed first pass in updating DGIMEXscheme.tex
Added text for the Fermion Implosion problem
Changed section title
Edits througout after reading the entire paper.

\section{Realizability-Preserving Limiter}
\label{eq:limiter}

The bound-preserving DG-IMEX method developed in previous sections is designed to preserve realizability of the cell averaged moments, i.e., $\vect{\cM}_{\bK}\in\cR$, provided sufficiently accurate quadratures are used to integrate integrals in the DG method, specific CFL conditions are satisfied, and that the polynomial approximation $\vect{\cM}_{h}$, at time $t^{n}$, is realizable in a set of quadrature points in each element $\bK$.  
We denote this quadrature set by $S\subset\bK$.  
In the DG method, we use the limiter proposed by Zhang \& Shu \cite{zhangShu_2010a} for scalar conservation laws to enforce the bounds on the zeroth moment $\cJ$ (see also \cite{liuOsher_1996}).  
We replace the polynomial $\cJ_{h}^{n}(\vect{z})$ with the limited polynomial
\begin{equation}
  \tilde{\cJ}_{h}^{n}(\vect{z})
  =\vartheta_{1}\,\cJ_{h}^{n}(\vect{z})+(1-\vartheta_{1})\,\cJ_{\bK}^{n},
\end{equation}
where the limiter parameter $\vartheta_{1}$ is given by
\begin{equation}
  \vartheta_{1}
  =\min\Big\{\,\Big|\f{M-\cJ_{\bK}^{n}}{M_{S}-\cJ_{\bK}^{n}}\Big|,\Big|\f{m-\cJ_{\bK}^{n}}{m_{S}-\cJ_{\bK}^{n}}\Big|,1\,\Big\},
\end{equation}
with $m=0$ and $M=1$, and
\begin{equation}
  M_{S}=\max_{\vect{z}\in S}\cJ_{h}^{n}(\vect{z})
  \quad\text{and}\quad
  m_{S}=\min_{\vect{z}\in S}\cJ_{h}^{n}(\vect{z}).  
\end{equation}

In the next step, we ensure realizability of the moments by following the framework of \cite{zhangShu_2010b}, developed to ensure positivity of the pressure when solving the Euler equations of gas dynamics.  
We let $\widetilde{\vect{\cM}}_{h}^{n}=\big(\tilde{\cJ}_{h}^{n},\vect{\cH}_{h}^{n}\big)^{T}$.  
Then, if $\widetilde{\bcM}_{h}^{n}$ lies outside $\cR$ for any quadrature point $\vect{z}_{q}\in S$, i.e., $\gamma(\widetilde{\bcM}_{h}^{n})<0$, there exists an intersection point of the straight line, $\vect{s}_{q}(\xi)$, connecting $\vect{\cM}_{\bK}^{n}\in\cR$ and $\widetilde{\vect{\cM}}_{h}^{n}$ evaluated in the troubled quadrature point $\vect{z}_{q}$, denoted $\widetilde{\vect{\cM}}_{q}^{n}$, and the boundary of $\cR$.  
This line is given by the convex combination 
\begin{equation}
  \vect{s}_{q}(\xi)=\xi\,\widetilde{\vect{\cM}}_{q}^{n}+(1-\xi)\,\bcM_{\bK}^{n},
\end{equation}
where $\xi\in[0,1]$, and the intersection point $\xi_{q}$ is obtained by solving $\gamma(\bs_{q}(\xi))=0$ for $\xi$, using the bisection algorithm.  
We then replace the polynomial representation $\widetilde{\vect{\cM}}_{h}^{n}\to\widehat{\vect{\cM}}_{h}^{n}$, where
\begin{equation}
  \widehat{\vect{\cM}}_{h}^{n}(\vect{z})=\vartheta_{2}\,\widetilde{\vect{\cM}}_{h}^{n}(\vect{z})+(1-\vartheta_{2})\,\vect{\cM}_{\bK}^{n},
\end{equation}
and $\vartheta_{2}=\min_{q}\xi_{q}$ is the smallest $\xi$ obtained in the element by considering all the troubled quadrature points.

We note that the realizability-preserving limiter is conservative; i.e., preserves the cell-average.

>>>>

\section{Realizability-Enforcing Limiter}
\label{sec:limiter}

The bound-preserving DG-IMEX method developed in previous sections is designed to preserve realizability of the cell averaged moments, i.e., $\vect{\cM}_{\bK}\in\cR$, provided sufficiently accurate quadratures are used to integrate integrals in the DG method, a CFL condition is satisfied, and that the polynomial approximation $\vect{\cM}_{h}$, at time $t^{n}$, is realizable in a set of quadrature points in each element $\bK$.  
We denote this quadrature set by $S=\cup_{k=1}^{d}\hat{\vect{S}}^{k}\subset\bK$.  
In the DG method, we use the limiter proposed by Zhang \& Shu \cite{zhangShu_2010a} for scalar conservation laws to enforce the bounds on the zeroth moment $\cJ$ (see also \cite{liuOsher_1996}).  
We replace the polynomial $\cJ_{h}^{n}(\vect{x})$ with the limited polynomial
\begin{equation}
  \tilde{\cJ}_{h}^{n}(\vect{x})
  =\vartheta_{1}\,\cJ_{h}^{n}(\vect{x})+(1-\vartheta_{1})\,\cJ_{\bK}^{n},
  \label{eq:limitDensity}
\end{equation}
where the limiter parameter $\vartheta_{1}$ is given by
\begin{equation}
  \vartheta_{1}
  =\min\Big\{\,\Big|\f{M-\cJ_{\bK}^{n}}{M_{S}-\cJ_{\bK}^{n}}\Big|,\Big|\f{m-\cJ_{\bK}^{n}}{m_{S}-\cJ_{\bK}^{n}}\Big|,1\,\Big\},
\end{equation}
with $m=0$ and $M=1$, and
\begin{equation}
  M_{S}=\max_{\vect{x}\in S}\cJ_{h}^{n}(\vect{x})
  \quad\text{and}\quad
  m_{S}=\min_{\vect{x}\in S}\cJ_{h}^{n}(\vect{x}).  
\end{equation}

In the next step, we ensure realizability of the moments by following the framework of \cite{zhangShu_2010b}, developed to ensure positivity of the pressure when solving the Euler equations of gas dynamics.  
We let $\widetilde{\vect{\cM}}_{h}^{n}=\big(\tilde{\cJ}_{h}^{n},\vect{\cH}_{h}^{n}\big)^{T}$.  
Then, if $\widetilde{\bcM}_{h}^{n}$ lies outside $\cR$ for any quadrature point $\vect{x}_{q}\in S$, i.e., $\gamma(\widetilde{\bcM}_{h}^{n})<0$, there exists an intersection point of the straight line, $\vect{s}_{q}(\xi)$, connecting $\vect{\cM}_{\bK}^{n}\in\cR$ and $\widetilde{\vect{\cM}}_{h}^{n}$ evaluated in the troubled quadrature point $\vect{x}_{q}$, denoted $\widetilde{\vect{\cM}}_{q}^{n}$, and the boundary of $\cR$.  
This line is given by the convex combination 
\begin{equation}
  \vect{s}_{q}(\xi)=\xi\,\widetilde{\vect{\cM}}_{q}^{n}+(1-\xi)\,\bcM_{\bK}^{n},
\end{equation}
where $\xi\in[0,1]$, and the intersection point $\xi_{q}$ is obtained by solving $\gamma(\bs_{q}(\xi))=0$ for $\xi$, using the bisection algorithm\footnote{In practice, $\xi$ needs not be accurate to many significant digits, and the bisection algorithm can be terminated after a few iterations.}.  
We then replace the polynomial representation $\widetilde{\vect{\cM}}_{h}^{n}\to\widehat{\vect{\cM}}_{h}^{n}$, where
\begin{equation}
  \widehat{\vect{\cM}}_{h}^{n}(\vect{x})=\vartheta_{2}\,\widetilde{\vect{\cM}}_{h}^{n}(\vect{x})+(1-\vartheta_{2})\,\vect{\cM}_{\bK}^{n},
  \label{eq:limitMoments}
\end{equation}
and $\vartheta_{2}=\min_{q}\xi_{q}$ is the smallest $\xi$ obtained in the element by considering all the troubled quadrature points.

We note that the realizability-enforcing limiter is conservative; i.e., preserves the cell-average $\widehat{\vect{\cM}}_{\bK}=\widetilde{\vect{\cM}}_{\bK}=\vect{\cM}_{\bK}$.
